---
title: "Introduction to data wrangling with `dplyr`"
author: "Stephanie J. Spielman"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(learnr)
tutorial_options(exercise.lines = 6)
```


# CB2R Data Science Workshop, Summer 2020
<br><br><br>

> The library `dplyr` has been pre-loaded for use using the command `library(dplyr)`.


### Piping syntax

```{r, eval=FALSE}
# Without piping
dplyr_function( data_frame_variable, arguments )

# With piping
data_frame_variable %>%
  dplyr_function( arguments )
```
<br><br>

## Set 1: Using dplyr verbs 

<br><br>

#### 1. Use `filter()` to subset `msleep` to only herbivores.
```{r dplyr1_q1, exercise = TRUE}

```

<br>


#### 2. Use `select()` to subset `msleep` to keep only the columns `name`, `awake`, `sleep_total`, `sleep_rem`, and `sleep_cycle`

```{r dplyr1_q2, exercise = TRUE}

```

<br>


#### 3. Use `select()` to subset `msleep` to *remove* the columns `genus` and `order`

```{r dplyr1_q3, exercise = TRUE}

```


<br>

#### 4. Use `filter()` to subset `msleep` to only animals who are *awake* for at least 12 hours of the day
 
```{r dplyr1_q4, exercise = TRUE}

```

<div id= "dplyr1_q4-hint">
**Hint:** "At least" like saying "greater than or equal to," i.e. `>=`
</div>  

<br>

#### 5. Use `filter()` to subset `msleep` to only *herbivores* who are *awake* for at least 12 hours of the days.
  
```{r dplyr1_q5, exercise = TRUE}

```  

<div id= "dplyr1_q5-hint">
**Hint:** Supply multiple "and" conditions to `filter()` with commas, i.e. `filter(statement1, statement2)`
</div> 

<br>

#### 6. Use `filter()` to subset `msleep` to only *herbivores OR insectivores* who are *awake* for at least 12 hours of the days.

```{r dplyr1_q6, exercise = TRUE}

```   

<div id= "dplyr1_q6-hint">
**Hint:** Supply multiple "and" conditions to `filter()` with commas, i.e. `filter(statement1, statement2)`
</div>
  
<br>

#### 7. Use `arrange()` to sort the dataset in *ascending* order of `bodywt`

```{r dplyr1_7, exercise = TRUE}

``` 

<br>

#### 8. Use `arrange()` to sort the dataset in *descending* order of `bodywt`
 
```{r dplyr1_q8, exercise = TRUE}

``` 

<div id= "dplyr1_q8-hint">
**Hint:** User `desc()` to sort by descending order of a column, like: `arrange(desc(COLUMN))`
</div> 

<br>

#### 9. Use `mutate()` to create a new column called `percent_day_awake` that gives the percentage of the day that each species spends awake. 
  
```{r dplyr1_q9, exercise = TRUE}

``` 

<div id= "dplyr1_q9-hint">
**Hint:** There are 24 hours in a day, and the column `awake` says how many hours a day (on average) that species is awake.
</div> 
 
<br>

#### 10. Use `mutate()` to create a new column called `log_bodywt` that gives the natural logarithm of the bodyweight.
  
```{r dplyr1_q10, exercise = TRUE}

``` 

<div id= "dplyr1_q10-hint">
**Hint:** `log()` by default calculates the natural logarithm
</div> 

<br>

#### 11. Use `rename()` to change the name of the column `conservation` to `conservation_status`.

```{r dplyr1_q11, exercise = TRUE}

``` 

## Set 2: Creating dplyr pipelines

<!-- Name these dplyr1.2_q1-... -->


<br>

#### 1. Use `filter()` to subset `msleep` to only herbivores, and then use `arrange()` to order the data by `name`. 

```{r dplyr1-2_q1, exercise = TRUE}

``` 

<br>

#### 2. Use `filter()` to subset `msleep` to only species whose conservation status is least concern ("lc"), and then use `select()` to remove the `conservation` column. **Save the final output** of your piped commands to a new data frame called `msleep_lc`, and then print the new data frame to confirm your work was successful.

```{r dplyr1-2_q2, exercise = TRUE}

``` 

<br>

#### 3. Use `filter()` to subset `msleep` to only primate species (order is "Primates") whose conservation status is least concern ("lc") (two things to filter!!), use `select()` to remove BOTH the `conservation` and `order` columns, and rename (using `rename()`) the column `vore` to be called `diet`.

```{r dplyr1-2_q3, exercise = TRUE}

``` 

<br>

#### 4. We can also pipe the output directly into other functions that aren't specifically part of the `dplyr` package. Copy/paste your output from question 3 below, and pipe the output into the function `nrow()` - this will tell you how many rows (species!!) are in the output, aka how many primates are of "least concern" status.

```{r dplyr1-2_q4, exercise = TRUE}

``` 

## Answer Keys


<br><br>

### Set 1 Answer Key

<br><br>

#### 1. Use `filter()` to subset `msleep` to only herbivores.

``` {r, dplyr1_q1-answer}
#Solution to Question 1
msleep %>%
  filter(vore == "herbi")
```

<br>

#### 2. Use `select()` to subset `msleep` to keep only the columns `name`, `awake`, `sleep_total`, `sleep_rem`, and `sleep_cycle`

``` {r, dplyr1_q2-answer}
#Solution to Question 2
msleep %>%
  select(name, awake, sleep_total, sleep_rem, sleep_cycle)
```

<br>

#### 3. Use `select()` to subset `msleep` to *remove* the columns `genus` and `order`

``` {r, dplyr1_q3-answer}
#Solution to Question 3
msleep %>%
  select(-genus, -order)
```

<br>

#### 4. Use `filter()` to subset `msleep` to only animals who are *awake* for at least 12 hours of the day
  
``` {r, dplyr1_q4-answer}
#Solution to Question 4
msleep %>%
  filter(awake >= 12)
```

<br>

#### 5. Use `filter()` to subset `msleep` to only *herbivores* who are *awake* for at least 12 hours of the days.
 
``` {r, dplyr1_q5-answer}
#Solution to Question 5
msleep %>%
  filter(vore == "herbi", awake >= 12)
```
  
<br>

#### 6. Use `filter()` to subset `msleep` to only *herbivores OR insectivores* who are *awake* for at least 12 hours of the days.
  
``` {r, dplyr1_q6-answer}
#Solution to Question 6
msleep %>%
  filter(vore %in% c("herbi", "insecti"), awake >= 12)
```
  
<br>

#### 7. Use `arrange()` to sort the dataset in *ascending* order of `bodywt`

``` {r, dplyr1_q7-answer}
#Solution to Question 7
msleep %>%
  arrange(bodywt)
```

<br>

#### 8. Use `arrange()` to sort the dataset in *descending* order of `bodywt`
 
``` {r, dplyr1_q8-answer}
#Solution to Question 8
msleep %>%
  arrange(desc(bodywt))
```  
  
<br>

#### 9. Use `mutate()` to create a new column called `percent_day_awake` that gives the percentage of the day that each species spends awake. 
  
``` {r, dplyr1_q9-answer}
#Solution to Question 9
msleep %>%
  mutate(percent_day_awake = (awake/24)*100)
```  
  
<br>

#### 10. Use `mutate()` to create a new column called `log_bodywt` that gives the natural logarithm of the bodyweight.
  
``` {r, dplyr1_q10-answer}
#Solution to Question 10
msleep %>%
  mutate(log_bodywt = log(bodywt))
```    

<br>

#### 11. Use `rename()` to change the name of the column `conservation` to `conservation_status`.

``` {r, dplyr1_q11-answer}
#Solution to Question 11
msleep %>%
  rename(conservation_status = conservation)
``` 

<br><br><br>

### Set 2 Answer Key

<br><br>

#### 1. Use `filter()` to subset `msleep` to only herbivores, and then use `arrange()` to order the data by `name`. 

``` {r, dplyr1-2_q1-answer}
#Solution to Question 1
msleep %>%
  filter(vore == "herbi") %>%
  arrange(name)
``` 

<br>

#### 2. Use `filter()` to subset `msleep` to only species whose conservation status is least concern ("lc"), and then use `select()` to remove the `conservation` column. **Save the final output** of your piped commands to a new data frame called `msleep_lc`, and then print the new data frame to confirm your work was successful.

``` {r, dplyr1-2_q2-answer}
#Solution to Question 2
msleep %>%
  filter(conservation == "lc") %>%
  select(-conservation) -> msleep_lc

msleep_lc
``` 

<br>

#### 3. Use `filter()` to subset `msleep` to only primate species (order is "Primates") whose conservation status is least concern ("lc") (two things to filter!!), use `select()` to remove BOTH the `conservation` and `order` columns, and rename (using `rename()`) the column `vore` to be called `diet`.

``` {r, dplyr1-2_q3-answer}
#Solution to Question 3
msleep %>%
  filter(order == "Primates", conservation == "lc") %>%
  select(-conservation, -order) %>%
  rename(diet = vore)
``` 

<br>

#### 4. We can also pipe the output directly into other functions that aren't specifically part of the `dplyr` package. Copy/paste your output from question 3 below, and pipe the output into the function `nrow()` - this will tell you how many rows (species!!) are in the output, aka how many primates are of "least concern" status. 

``` {r, dplyr1-2_q4-answer}
#Solution to Question 4
msleep %>%
  filter(order == "Primates", conservation == "lc") %>%
  select(-conservation, -order) %>%
  rename(diet = vore) %>%
  nrow()
``` 

